---
title: "MetaQCAT: microbiome meta-analysis across studies"
author: "Shuren He"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    css: metaqcat.css
    fig_caption: true
    mathjax: null
vignette: >
  %\VignetteIndexEntry{MetaQCAT: microbiome meta-analysis across studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 4.5,
  fig.align = "center", dpi = 140, message = FALSE,
  error = FALSE, cache = FALSE
)
options(width = 88)
library(MetaQCAT)
data("CRC_data", package = "MetaQCAT")
data("IBD_data", package = "MetaQCAT")
```

<div class="lead">
A practical guide to composition association tests, study-specific covariates,
and subject-level clustering, with the packaged CRC and IBD datasets.
</div>

<div class="workflow">
<div><b>01 · Prepare</b><span>Counts, covariates, taxonomy, and sample alignment</span></div>
<div><b>02 · Choose</b><span>Fixed or random effects; composition or zero patterns</span></div>
<div><b>03 · Fit</b><span>Independent samples or within-study clusters</span></div>
<div><b>04 · Report</b><span>Lineage p-values, multiplicity, and reproducibility</span></div>
</div>

# What MetaQCAT tests

`MetaQCAT()` combines microbiome association information across studies.
This release is built from MetaPALM, with consistent argument names and
checked R/C++ interfaces. Its input is a list of **sample-by-taxon count matrices**, accompanied by
covariate matrices and, when needed, subject identifiers. A taxonomy matrix
defines the lineages to be tested.

For example, a family-to-genus taxonomy asks whether the **composition of
genera within each family** is associated with disease. This is not the same
question as testing the total abundance of a family relative to the whole
microbiome. Interpret the result at the level of the composition being tested.

| Method | Information used | Calibration |
|:--|:--|:--|
| `FE-MetaQCAT` | Composition; common-effect component | Asymptotic or resampling |
| `FE-MetaQCAT*` | Zero patterns | Asymptotic or resampling |
| `FE-MetaQCAT-O` | Cauchy combination of the FE components | Asymptotic or resampling |
| `RE-MetaQCAT` | Composition plus an effect-heterogeneity component | Resampling required |
| `RE-MetaQCAT*` | Zero-pattern random-effect component | Resampling required |
| `RE-MetaQCAT-O` | Cauchy combination of the RE components | Resampling required |

The choice of fixed versus random effects should reflect the scientific
question and plausible cross-study effect heterogeneity. A significant RE
result is an association result, not a separate proof of heterogeneity. The
omnibus methods combine p-values; they do not average coefficient estimates.

<div class="callout">
<p><strong>Which implementation?</strong> This vignette uses the MetaPALM-based
release. Its zero-pattern resampling is connected to the exported C++ backend.
All six methods are exercised below. Successful software execution is not a
proof of type-I error control: calibration and exchangeability should also be
assessed for the intended study design.</p>
</div>

# The two example datasets

```{r load-data}
library(MetaQCAT)
data("CRC_data", package = "MetaQCAT")
data("IBD_data", package = "MetaQCAT")

names(CRC_data)
names(IBD_data)
```

Both `.rda` files store one named list. Common field names are consistent,
but the datasets retain their actual study designs rather than inventing
covariates or clusters merely to make their structures identical.

| Field | CRC | IBD |
|:--|:--|:--|
| `counts` | Five count matrices | Ten count matrices |
| `metadata` | Original CRC sample metadata | Original IBD metadata and derived study labels |
| `disease` | One column: CRC = 1, control = 0 | One column: CD/UC = 1, control = 0 |
| `design` | Not stored; use `disease` directly | Disease followed by study-specific covariates |
| `subject_ids` | `NULL`; independent-sample analysis | Subject factors for repeated samples |
| `taxonomy` | Original numeric-prefix annotations | Original `k__` through `g__` annotations |
| `adjust` | Zero-column matrices | Adjustment columns of `design` |

CRC contains 573 samples (283 cases and 290 controls). IBD contains 3,472
samples in ten analysis strata. HMP2 is split at age 18; MucosalIBD is retained
as one study. The package data are already prevalence-filtered within each
study. See `?CRC_data` and `?IBD_data` for exact preprocessing and covariate
coding, including how missing values were handled.

## Summarize study sizes

```{r study-summary}
study_summary <- function(d) {
  data.frame(
    study = names(d$counts),
    samples = vapply(d$counts, nrow, integer(1)),
    cases = vapply(d$disease, function(x) sum(x[, 1] == 1), integer(1)),
    controls = vapply(d$disease, function(x) sum(x[, 1] == 0), integer(1)),
    taxa = vapply(d$counts, ncol, integer(1)),
    row.names = NULL
  )
}
crc_summary <- study_summary(CRC_data)
ibd_summary <- study_summary(IBD_data)
knitr::kable(crc_summary, caption = "CRC studies after preprocessing")
knitr::kable(ibd_summary, caption = "IBD analysis strata after preprocessing")
```

```{r cohort-figure, echo=FALSE, fig.height=6, fig.cap="Case/control sample counts by analysis study. Repeated IBD samples are not independent participants."}
local({
  old <- par(no.readonly = TRUE)
  on.exit({ layout(1); par(old) })
  layout(matrix(c(1, 2, 3, 3), nrow = 2, byrow = TRUE), heights = c(9, 1))
  par(las = 1, family = "sans", cex.main = 1,
      col.axis = "#465b68", col.lab = "#213342")
  draw <- function(s, title, labels) {
    values <- rbind(Control = s$controls, Case = s$cases)
    barplot(values[, rev(seq_len(ncol(values))), drop = FALSE],
            horiz = TRUE, names.arg = rev(labels), border = NA,
            col = c("#88b9c6", "#087f8c"), cex.names = 0.75,
            main = title, xlab = "Samples", space = 0.5)
  }
  par(mar = c(4, 4, 3, 1))
  draw(crc_summary, "CRC · independent samples", crc_summary$study)
  par(mar = c(4, 10, 3, 1))
  labels <- c("CS-PRISM biopsy", "CS-PRISM stool", "Herfarth stool",
              "HMP2 adult biopsy", "HMP2 pediatric biopsy", "Jansson stool",
              "MucosalIBD biopsy", "Pouchitis biopsy", "RISK biopsy", "RISK stool")
  draw(ibd_summary, "IBD · repeated samples", labels)
  par(mar = c(0, 0, 0, 0))
  plot.new()
  legend("center", c("Control", "Case"), fill = c("#88b9c6", "#087f8c"),
         border = NA, bty = "n", horiz = TRUE, cex = 0.9)
})
```

# Prepare inputs correctly

## Counts and covariates

Supply raw count matrices, not percentages or log-transformed abundances.
Rows of `counts[[i]]` and `X[[i]]` must refer to the same samples in the same
order. Taxon columns may differ between studies, but their names must be
unique and covered by the taxonomy. Do not replace missing measurements with
zeros merely to satisfy input requirements.

`X` is a conventional statistical design-matrix name. **Do not add an
intercept**: MetaQCAT adds it internally. `X.index` refers to your input
columns, before that intercept is added.

```{r input-checks}
for (s in names(CRC_data$counts)) {
  stopifnot(
    identical(rownames(CRC_data$counts[[s]]),
              rownames(CRC_data$disease[[s]])),
    all(colnames(CRC_data$counts[[s]]) %in% rownames(CRC_data$taxonomy)),
    all(is.finite(CRC_data$counts[[s]])),
    !anyNA(CRC_data$disease[[s]])
  )
}

colnames(IBD_data$design[["HMP2.adult.biopsy"]])
colnames(IBD_data$design[["MucosalIBD.biopsy"]])
```

With `X = IBD_data$design`, `X.index = 1` tests disease while retaining
the remaining columns as adjustment covariates. To test several variables
jointly, supply a vector of positions, for example `c(1, 2)`, only when those
positions have the same scientific meaning in every study. Omission of
`X.index` retains the existing behavior of testing all input columns; specify
it explicitly for adjusted analyses.

## Define a family-to-genus taxonomy

```{r family-taxonomy}
crc_tax <- CRC_data$taxonomy[, c("Rank5", "Rank6"), drop = FALSE]
colnames(crc_tax) <- c("Rank1", "Rank2")

ibd_tax <- IBD_data$taxonomy[, c("Rank5", "Rank6"), drop = FALSE]
colnames(ibd_tax) <- c("Rank1", "Rank2")
```

Here `Rank1` means family and `Rank2` means genus in the **analysis taxonomy**.
The rank numbering is relative to the supplied matrix; it need not start at
kingdom. Families with only one eligible child genus cannot provide a
within-family composition contrast. Higher-rank analyses can use more columns
of the original taxonomy, in higher-to-lower order.

## A small, explicitly filtered teaching subset

The helper below restricts the demonstration to selected families and removes
samples with zero total count in that restricted composition. It subsets
counts, covariates, and subject IDs together, avoiding sample misalignment.
It is a vignette helper, not another exported package function.

```{r prepare-helper}
prepare_lineages <- function(d, families, studies = names(d$counts),
                             adjusted = FALSE) {
  tax <- d$taxonomy[d$taxonomy[, "Rank5"] %in% families,
                    c("Rank5", "Rank6"), drop = FALSE]
  stopifnot(nrow(tax) > 1, all(studies %in% names(d$counts)))
  colnames(tax) <- c("Rank1", "Rank2")
  X <- if (adjusted) d$design[studies] else d$disease[studies]
  stopifnot(length(X) == length(studies))
  counts <- d$counts[studies]
  ids <- if (is.null(d$subject_ids)) NULL else d$subject_ids[studies]
  removed <- setNames(integer(length(studies)), studies)
  for (s in studies) {
    y <- counts[[s]][, intersect(colnames(counts[[s]]), rownames(tax)),
                      drop = FALSE]
    keep <- rowSums(y) > 0
    stopifnot(any(keep), ncol(y) > 1)
    removed[s] <- sum(!keep)
    counts[[s]] <- y[keep, , drop = FALSE]
    X[[s]] <- X[[s]][keep, , drop = FALSE]
    if (!is.null(ids)) ids[[s]] <- factor(ids[[s]][keep])
    stopifnot(identical(rownames(counts[[s]]), rownames(X[[s]])))
  }
  list(counts = counts, X = X, cluster.id = ids,
       taxonomy = tax, removed = removed)
}
```

<div class="callout">
<p><strong>Demonstration, not a published reanalysis.</strong> Restricting the
composition changes sample eligibility and, for the zero-part tests, the
within-study rarefaction depth. The displayed p-values illustrate the workflow;
they are not substitutes for a prespecified analysis of the full data.</p>
</div>

# CRC: independent-sample analysis

## Fit the composition test

```{r crc-fit}
crc_demo <- prepare_lineages(
  CRC_data,
  families = c("543 Enterobacteriaceae", "186804 Peptostreptococcaceae")
)
crc_demo$removed

crc_fe <- MetaQCAT(
  counts = crc_demo$counts,
  X = crc_demo$X,
  X.index = 1,
  method = "FE-MetaQCAT",
  taxonomy = crc_demo$taxonomy,
  min.depth = 0,
  min.prev = 0.1,
  save.coef = TRUE,
  seed = 1234,
  n.threads = 1
)
stopifnot(length(crc_fe$failed.lineage) == 0L)
knitr::kable(t(crc_fe$lineage.pval), digits = 5,
             caption = "Illustrative FE composition tests")
```

CRC has no separate stored `design`: the one-column disease indicator is the
complete input `X` for this unadjusted example. Independent samples do not
require `cluster.id` or a clustered permutation scheme.

## Resampling and random effects

```{r crc-resampling}
crc_fe_perm <- MetaQCAT(
  counts = crc_demo$counts, X = crc_demo$X, X.index = 1,
  method = "FE-MetaQCAT", taxonomy = crc_demo$taxonomy,
  n.perm = 99, n.threads = 1
)
crc_re <- MetaQCAT(
  counts = crc_demo$counts, X = crc_demo$X, X.index = 1,
  method = "RE-MetaQCAT", taxonomy = crc_demo$taxonomy,
  n.perm = 99, n.threads = 1
)
knitr::kable(t(crc_fe_perm$lineage.pval), digits = 4)
knitr::kable(t(crc_re$lineage.pval), digits = 4)
```

The 99-resample examples keep the vignette fast. They are too coarse for
fine-grained inference or substantial multiple testing. The implementation
uses adaptive resampling, so `n.perm` is a requested budget rather than a
guarantee that every lineage performs exactly that number of draws. Larger
budgets should be chosen before analysis according to the resolution needed.
Random-effect methods require a non-`NULL` `n.perm`; no asymptotic RE p-value
is returned by this interface.

## Zero patterns and the asymptotic omnibus test

```{r crc-zero-omnibus}
crc_zero <- MetaQCAT(
  counts = crc_demo$counts, X = crc_demo$X, X.index = 1,
  method = "FE-MetaQCAT*", taxonomy = crc_demo$taxonomy,
  n.perm = NULL, n.threads = 1
)
crc_omnibus <- MetaQCAT(
  counts = crc_demo$counts, X = crc_demo$X, X.index = 1,
  method = "FE-MetaQCAT-O", taxonomy = crc_demo$taxonomy,
  n.perm = NULL, n.threads = 1
)
knitr::kable(t(crc_zero$lineage.pval), digits = 5)
knitr::kable(t(crc_omnibus$lineage.pval), digits = 5)
```

Starred tests internally rarefy each study to its minimum retained sample
depth before constructing zero indicators. Consequently, their zero patterns
can differ from zeros in the original counts. Very low-depth samples can
therefore influence the zero-part analysis even when they have nonzero depth.
By default, the zero component retains taxa with within-study zero proportions
strictly between 0.1 and 0.9. This is a two-sided variability filter. For
composition tests, `min.prev = 0.1` retains taxa present in at least 10% of
samples. An omnibus test can use separate thresholds, e.g.
`min.prev = c(0.1, 0.05)`.

## Zero-pattern and omnibus resampling

The same interface supports the remaining resampling paths, including the
random-effect zero and omnibus methods.

```{r crc-all-methods}
methods <- c("FE-MetaQCAT*", "FE-MetaQCAT-O",
             "RE-MetaQCAT*", "RE-MetaQCAT-O")
crc_other <- setNames(lapply(methods, function(m) {
  fit <- MetaQCAT(
    counts = crc_demo$counts, X = crc_demo$X, X.index = 1,
    taxonomy = crc_demo$taxonomy, method = m,
    min.prev = 0.1, n.perm = 99, seed = 1234
  )
  stopifnot(length(fit$failed.lineage) == 0L,
            all(is.finite(fit$lineage.pval)))
  fit
}), methods)
other_p <- do.call(rbind, lapply(crc_other, function(r)
  r$lineage.pval[nrow(r$lineage.pval), , drop = FALSE]))
knitr::kable(other_p, digits = 4,
             caption = "Zero-pattern and omnibus resampling examples")
```

All methods use the same study inputs; the zero component rarefies once per
study per call, rather than once per lineage. Sequential hypergeometric
sampling avoids expanding every sample into a vector of reads. It has the
same rarefaction distribution as the earlier read-expansion implementation,
although exact results for the same seed can differ.

# IBD: covariate adjustment and repeated samples

## Check subject overlap across analysis strata

Cluster identifiers account for dependence **within a study**. They do not
automatically account for the same participant contributing to multiple study
lists. The packaged IBD data include analysis strata by sample type and age,
and some participant identifiers occur in more than one stratum.

```{r overlap-check}
members <- do.call(rbind, lapply(names(IBD_data$subject_ids), function(s) {
  data.frame(subject = unique(as.character(IBD_data$subject_ids[[s]])),
             study = s)
}))
overlap <- table(members$subject)
sum(overlap > 1)
```

Review overlapping cohorts before a substantive meta-analysis. Do not treat
biopsy/stool strata from overlapping participants as independent studies merely
because they have different names. This example uses three biopsy cohorts
and explicitly verifies that their subject identifiers do not overlap.

## Set up cluster-aware inputs

```{r ibd-inputs}
ibd_studies <- c("HMP2.adult.biopsy", "MucosalIBD.biopsy", "RISK.biopsy")
selected_members <- members[members$study %in% ibd_studies, ]
stopifnot(!anyDuplicated(selected_members$subject))

ibd_demo <- prepare_lineages(
  IBD_data, families = "f__Neisseriaceae",
  studies = ibd_studies, adjusted = TRUE
)
ibd_demo$removed
lapply(ibd_demo$X, colnames)
```

`X` contains disease plus the available adjustment covariates; `X.index = 1`
targets disease only. `cluster.id` contains factors based on real subject IDs,
not the covariate matrices. The two objects have different roles and must not
be interchanged.

## Choose the permutation unit

| Scheme | Appropriate structure | Main check |
|:--|:--|:--|
| `"between"` | Tested exposure is constant within each subject/cluster | Cluster-level exchangeability is defensible |
| `"within"` | Tested exposure varies within subject/cluster | Within-cluster exchanges are defensible |

Disease is constant within subjects in this example, so a within-subject shuffle
would not change disease labels. We use `"between"`. Covariate adjustment does
not itself guarantee exchangeability; the permutation scheme must still match
the study design and scientific null hypothesis.

```{r cluster-label-check}
for (s in names(ibd_demo$X)) {
  by_subject <- split(ibd_demo$X[[s]][, "disease"], ibd_demo$cluster.id[[s]])
  stopifnot(all(vapply(by_subject, function(x) length(unique(x)) == 1,
                       logical(1))))
}
```

## Fit FE and RE composition tests

```{r ibd-fits}
ibd_fe <- MetaQCAT(
  counts = ibd_demo$counts, X = ibd_demo$X, X.index = 1,
  method = "FE-MetaQCAT", taxonomy = ibd_demo$taxonomy,
  cluster.id = ibd_demo$cluster.id, permute.strata = "between",
  n.perm = 99, n.threads = 1
)
ibd_re <- MetaQCAT(
  counts = ibd_demo$counts, X = ibd_demo$X, X.index = 1,
  method = "RE-MetaQCAT", taxonomy = ibd_demo$taxonomy,
  cluster.id = ibd_demo$cluster.id, permute.strata = "between",
  n.perm = 99, n.threads = 1
)
knitr::kable(t(ibd_fe$lineage.pval), digits = 4)
knitr::kable(t(ibd_re$lineage.pval), digits = 4)
```

This is a small instructional subset with low resampling resolution. Interpret
neither the displayed FE/RE difference nor the 0.05 threshold as a confirmed
biological finding. A final analysis needs prespecified study inclusion,
covariate handling, adequate calibration, and sensitivity checks.

# Read, adjust, and visualize results

## Result orientation and BH adjustment

Rows of `lineage.pval` identify calibration/method and columns identify
lineages. Use `drop = FALSE` when retaining a matrix with one row or column.
The package's `sig.lineage` applies BH adjustment to returned, nonmissing
lineage p-values: resampling p-values when requested, otherwise asymptotic ones.
Failed tests retain named `NA` columns and reasons in `failed.lineage`. The
BH denominator includes all attempted eligible lineages, including failures;
failed tests are never declared significant. `lineage.qval` reports the adjusted
values directly.

```{r results-table}
p <- crc_fe_perm$lineage.pval["Resampling-FE-MetaQCAT", ]
results <- data.frame(
  lineage = names(p),
  p.value = as.numeric(p),
  q.value = p.adjust(p, method = "BH"),
  row.names = NULL
)
results <- results[order(results$p.value), ]
knitr::kable(results, digits = 4)
crc_fe_perm$sig.lineage
```

Adjust over the intended testing family, not just a hand-selected subset of
interesting results. BH-adjusting these two demonstration families is not the
same as adjusting all families in a full analysis. Repeatedly trying methods,
rank levels, or preprocessing thresholds also changes the inferential context.

```{r pvalue-figure, echo=FALSE, fig.height=3.8, fig.cap="Illustrative composition p-values for the two CRC teaching lineages. The dashed line is a nominal p = 0.05 reference, not an FDR threshold."}
local({
  v <- rbind(
    FE = crc_fe_perm$lineage.pval["Resampling-FE-MetaQCAT", ],
    RE = crc_re$lineage.pval["Resampling-RE-MetaQCAT", ]
  )
  old <- par(mar = c(5, 5, 2, 1), las = 1, family = "sans")
  on.exit(par(old))
  barplot(-log10(pmax(v, .Machine$double.xmin)), beside = TRUE,
          names.arg = sub("^[0-9]+ ", "", colnames(v)),
          col = c("#087f8c", "#7896ad"), border = NA,
          ylab = expression(-log[10](p)), ylim = c(0, 2.4), cex.names = 0.85)
  abline(h = -log10(0.05), lty = 2, col = "#b37a20")
  legend("topright", c("FE", "RE"), fill = c("#087f8c", "#7896ad"),
         border = NA, bty = "n", horiz = TRUE)
})
```

## Coefficients and reference taxa

```{r coefficient-structure}
str(crc_fe$lineage.coef[[1]], max.level = 2)
```

For the composition component, inspect `reference.taxon` in each lineage
result before interpreting `est.beta.comp`. Set `save.coef = TRUE` to retain
coefficients and `var.beta.comp` (variances, not standard errors). Estimates
are relative to the reference, not absolute taxon abundance changes. Take
square roots of variances only after checking they are finite and nonnegative.
The zero-part coefficients concern zero indicators and should not be
interpreted on the positive-part scale.

# Full analyses and reproducible reporting

The following template uses all CRC genera and the full family-to-genus
taxonomy. It is shown but **not executed** during vignette building, to avoid
running an expensive analysis whenever the package is installed or checked.

```{r full-analysis, eval=FALSE}
crc_full <- MetaQCAT(
  counts = CRC_data$counts,
  X = CRC_data$disease,
  X.index = 1,
  method = "FE-MetaQCAT",
  taxonomy = crc_tax,
  min.depth = 0,
  min.prev = 0.1,
  n.perm = 9999,
  fdr.alpha = 0.05,
  n.threads = 1
)

saveRDS(crc_full, file = "CRC_FE_results.rds")
write.csv(t(crc_full$lineage.pval), "CRC_FE_pvalues.csv")
capture.output(sessionInfo(), file = "CRC_sessionInfo.txt")
```

The resampling budget above is a template, not a universal recommendation.
Record dataset versions, exclusions, taxonomic ranks, tested variables,
adjustment sets, cluster definitions, permutation scheme, method, thresholds,
package/R versions, and resampling settings. Pass `seed` explicitly to control
rarefaction and resampling. Each lineage uses the supplied seed so that worker
scheduling does not change the result. To use independent random streams across
simulation replicates, provide different seeds for different calls.

For a single composition-wide test, omit taxonomy (`taxonomy = NULL`) and read
`fit$pvalue.comp`, `fit$pvalue.pa`, or `fit$pvalue.comb` rather than
`fit$lineage.pval` (depending on the selected method). Its scope is the entire supplied
composition, not a collection of lineage tests.

# Parameter reference and troubleshooting

## Release parameter names

```{r public-arguments}
args(MetaQCAT)
```

| Argument | Meaning / default |
|:--|:--|
| `counts` | Study-level count matrices |
| `X` | Study-level covariates; no intercept |
| `X.index` | Tested input-column indices; specify explicitly |
| `method` | `"FE-MetaQCAT"` by default |
| `cluster.id` | Subject/cluster IDs; `NULL` for independent samples |
| `taxonomy` | Taxonomic hierarchy; `NULL` for a composition-wide test |
| `min.depth` | Remove depths **less than or equal to** 0 by default |
| `min.prev` | Minimum prevalence, 0.1; two-sided variability threshold for zero tests |
| `n.perm` | `NULL`: asymptotic FE only; otherwise adaptive resampling budget |
| `permute.strata` | `NULL`, `"within"`, or `"between"`, according to design |
| `fdr.alpha` | BH threshold, 0.05 |
| `save.coef` | Store descendant coefficients/variances; FALSE |
| `n.threads` | One process by default |
| `seed` | Rarefaction/resampling seed; 1234 |

The release does not accept the old argument aliases.

| MetaPALM name | Release name |
|:--|:--|
| `OTU` | `counts` |
| `Tax` | `taxonomy` |
| `cluster` / `cluster.list` | `cluster.id` |
| `meta.method` | `method` |
| `prev.filter` | `min.prev` |
| `save.desc` | `save.coef` |
| `nthreads` | `n.threads` |
| `Y.list` / `X.list` / `X.par.index` | `counts` / `X` / `X.index` |
| `Y.bin.list` / `Z.list` / `Z.par.index` | `zeros` / `X` / `X.index` |

`X` and `X.index` retain familiar statistical notation. The six advanced
`Score_test_*` functions keep their names, but use the same concise arguments.
They require an intercept in column 1; unlike `MetaQCAT()`, they do not add one.
Zero-pattern advanced functions take binary zero indicators (1 = zero), not
raw counts. The default analysis uses the multinomial `Score_test_QCAT_*`
component; the retained Poisson `Score_test_*` variant is available for method
development and is not silently substituted into the routine analysis.

Lineage p-value and q-value matrices and failure diagnostics are now returned
explicitly. Component fields retain MetaPALM's `pvalue.comp`, `pvalue.pa`,
and `pvalue.comb` conventions; do not assume the result fields of the older
MetaQCAT implementation.

## Common pitfalls

| Symptom | What to check |
|:--|:--|
| Sample/covariate dimension mismatch | Study ordering and sample row names; subset every input together |
| Singular fit or `svd()` failure | Constant/redundant covariates, missing values, tiny eligible groups, or empty compositions |
| `class.id` / unused argument error | Use the release name `cluster.id`; never pass `design` as cluster IDs |
| Few or missing lineage results | Child-taxon eligibility, an available reference, fitting failures, and sparse data |
| NA columns | Inspect `failed.lineage`; failed tests retain their positions |
| A within-cluster shuffle has no effect | The tested exposure may be constant within every cluster; reconsider the permutation unit |
| Unexpected zero-part results | Internal rarefaction, low minimum depth, and zero-feature eligibility |
| Non-finite permutation statistic | The test fails explicitly; inspect sparse data or degenerate covariance, rather than reporting a p-value |
| Parallel execution fails on Windows | Use `n.threads = 1`; the backend is fork-based |

An unestimable lineage is not a nonsignificant finding. Do not replace failures
with `p = 1`, silently drop them from denominators, or interpret error-generated
values as inference. Preserve the reason for failure and distinguish omitted
tests from successfully tested lineages.

Routine depth filtering is quiet. Excluding an empty study or failing a
lineage produces a warning and diagnostics. Keep warnings visible and retain
`failed.lineage` with the exported results.

```{r help-pages, eval=FALSE}
help("MetaQCAT", package = "MetaQCAT")
help("CRC_data", package = "MetaQCAT")
help("IBD_data", package = "MetaQCAT")
citation("MetaQCAT")
```

## Session information

```{r session-info}
sessionInfo()
```
